---
title: "Chapter 5"
output: html_notebook
---

draw histograms and normal curves for scores on day1, 2 and 3

```{r}
dlf <- read.delim("../Data files/DownloadFestival(No Outlier).dat")
hist.day1 <- ggplot(dlf, aes(day1))
hist.day2 <- ggplot(dlf, aes(day2))
hist.day3 <- ggplot(dlf, aes(day3))

hist.day1 + geom_histogram(aes(y = ..density..), colour = "black", fill = "white") + stat_function(fun = dnorm, args = list(mean = mean(dlf$day1, na.rm =TRUE), sd = sd(dlf$day1, na.rm=TRUE)), colour = "black",size=1)
hist.day2 + geom_histogram(aes(y = ..density..), colour = "black", fill = "white")  + stat_function(fun = dnorm, args = list(mean = mean(dlf$day2, na.rm =TRUE), sd = sd(dlf$day2, na.rm=TRUE)), colour = "black",size=1)
hist.day3 + geom_histogram(aes(y = ..density..), colour = "black", fill = "white")  + stat_function(fun = dnorm, args = list(mean = mean(dlf$day3, na.rm =TRUE), sd = sd(dlf$day3, na.rm=TRUE)), colour = "black",size=1)
```

Test if day1, day2, and day3 scores are normally distributed. we compare against normally distributed data for all 3 days. if we have vertical line, it means that the data is perfectly normally distributed.

```{r}
qqplot.day1 <- ggplot(dlf, aes(sample=day1)) + stat_qq() + labs(x = "day1 theoretical score", y = "day1 actual score")
qqplot.day2 <- ggplot(dlf, aes(sample=day2)) + stat_qq() + labs(x = "day2 theoretical score", y = "day2 actual score")
qqplot.day3 <- ggplot(dlf, aes(sample=day3)) + stat_qq() + labs(x = "day3 theoretical score", y = "day3 actual score")
qqplot.day1
qqplot.day2
qqplot.day3
```

```{r}
describe(dlf$day1)
```

```{r}
stat.desc(dlf$day1, basic = FALSE, norm = TRUE)
```

```{r}
describe(cbind(dlf$day1, dlf$day2, dlf$day3))

```

```{r}
rexam <- read.delim("../Data files/rexam.dat", header=TRUE)
```

```{r}
str(rexam)
```

```{r}
rexam$uni <- factor(rexam$uni, levels = c(0:1), labels = c("Duncetown University", "Sussex University"))
```

```{r}
stat.desc(cbind(rexam$exam, rexam$computer, rexam$lectures, rexam$numeracy), basic = FALSE, norm = TRUE)
```

```{r}
exam_scores_plot <- ggplot(rexam, aes(exam))
computer_literacy_plot <- ggplot(rexam, aes(computer))
lectures_attended_plot <- ggplot(rexam, aes(lectures))
numeracy_attended_plot <- ggplot(rexam, aes(numeracy))

exam_scores_plot + geom_histogram(aes(y = ..density..), color="black", fill="white") + labs(x = "exam scrore") + stat_function(fun = dnorm, args = list(mean = mean(rexam$exam, na.rm =TRUE), sd = sd(rexam$exam, na.rm=TRUE)), colour = "blue",size=1)
computer_literacy_plot + geom_histogram(aes(y = ..density..), color="black", fill="white") + labs(x = "computer literacy") + stat_function(fun = dnorm, args = list(mean = mean(rexam$computer, na.rm =TRUE), sd = sd(rexam$computer, na.rm=TRUE)), colour = "blue",size=1)
lectures_attended_plot + geom_histogram(aes(y = ..density..), color="black", fill="white") + labs(x = "lectures attended") + stat_function(fun = dnorm, args = list(mean = mean(rexam$lectures, na.rm =TRUE), sd = sd(rexam$lectures, na.rm=TRUE)), colour = "blue",size=1)
numeracy_attended_plot + geom_histogram(aes(y = ..density..), color="black", fill="white") + labs(x = "numeracy test") + stat_function(fun = dnorm, args = list(mean = mean(rexam$numeracy, na.rm =TRUE), sd = sd(rexam$numeracy, na.rm=TRUE)), colour = "blue",size=1)
```

```{r}
by(cbind(data=rexam$exam,data=rexam$numeracy), rexam$uni, stat.desc, basic = FALSE, norm = TRUE)
dunceData<-subset(rexam, rexam$uni=="Duncetown University")
sussexData<-subset(rexam, rexam$uni=="Sussex University")
```

```{r}
hist.numeracy.duncetown <- ggplot(dunceData, aes(numeracy)) + geom_histogram(aes(y = ..density..), fill = "white",
colour = "black", binwidth=1) + labs(x = "Duncetown Numeracy Score", y = "Density") + stat_function(fun=dnorm, args=list(mean = mean(dunceData$numeracy, na.rm = TRUE), sd = sd(dunceData$numeracy, na.rm = TRUE)), colour = "blue",
size=1)
hist.numeracy.duncetown

hist.numeracy.sussex <- ggplot(sussexData, aes(numeracy)) + geom_histogram(aes(y = ..density..), fill = "white",
colour = "black", binwidth=1) + labs(x = "Sussex Numeracy Score", y = "Density") + stat_function(fun=dnorm, args=list(mean = mean(sussexData$numeracy, na.rm = TRUE), sd = sd(sussexData$numeracy, na.rm = TRUE)), colour = "blue",
size=1)
hist.numeracy.sussex
```

```{r}
by(cbind(data=rexam$lectures,data=rexam$computer), rexam$uni, stat.desc, basic = FALSE, norm = TRUE)
```

```{r}
hist.lectures.duncetown <- ggplot(dunceData, aes(lectures)) + geom_histogram(aes(y = ..density..), fill = "white",
colour = "black") + labs(x = "Duncetown lecture attendance percentage", y = "Density") + stat_function(fun=dnorm, args=list(mean = mean(dunceData$lectures, na.rm = TRUE), sd = sd(dunceData$lectures, na.rm = TRUE)), colour = "blue",
size=1)
hist.lectures.duncetown

hist.lectures.sussex <- ggplot(sussexData, aes(lectures)) + geom_histogram(aes(y = ..density..), fill = "white",
colour = "black") + labs(x = "Sussex lecture attendance percentage", y = "Density") + stat_function(fun=dnorm, args=list(mean = mean(sussexData$lectures, na.rm = TRUE), sd = sd(sussexData$lectures, na.rm = TRUE)), colour = "blue",
size=1)
hist.lectures.sussex
```

Q-Q plots for both exam and numeracy show that the data for these 2 variables are not normally distributed.

```{r}
shapiro.test(rexam$exam)
shapiro.test(rexam$numeracy)
```

```{r}
by(rexam$exam, rexam$uni, shapiro.test)
by(rexam$numeracy, rexam$uni, shapiro.test)
```

```{r}
ggplot(rexam, aes(sample=exam)) + stat_qq() + labs(x = "exam theoretical score", y = "exam actual score")

ggplot(rexam, aes(sample=numeracy)) + stat_qq() + labs(x = "numeracy theoretical score", y = "numeracy actual score")
```

Levene's test to test the equality of variance of exam scores and numeracy for each each university.

R exam score variances are similar for both universities

```{r}
leveneTest(rexam$exam, rexam$uni)
```

numeracy score variances are siginificantly different in 2 groups

```{r}
leveneTest(rexam$numeracy, rexam$uni)
```

Transform variables to fix the skewness of day2 and day3 scores. taking log of each values. 

```{r}
dlf$logday1 <- log(dlf$day1 + 1)
dlf$logday2 <- log(dlf$day2 + 1)
dlf$logday3 <- log(dlf$day3 + 1)

ggplot(dlf, aes(logday1)) + geom_histogram(aes(y = ..density..), fill = "white",
colour = "black")

ggplot(dlf, aes(logday2)) + geom_histogram(aes(y = ..density..), fill = "white",
colour = "black")

ggplot(dlf, aes(logday3)) + geom_histogram(aes(y = ..density..), fill = "white",
colour = "black")
```
```{r}
dlf$sqrtday1 <- sqrt(dlf$day1)
dlf$sqrtday2 <- sqrt(dlf$day2)
dlf$sqrtday3 <- sqrt(dlf$day3)

ggplot(dlf, aes(sqrtday1)) + geom_histogram(aes(y = ..density..), fill = "white",
colour = "black")

ggplot(dlf, aes(sqrtday2)) + geom_histogram(aes(y = ..density..), fill = "white",
colour = "black")

ggplot(dlf, aes(sqrtday3)) + geom_histogram(aes(y = ..density..), fill = "white",
colour = "black")
```

```{r}
# mean hygiene score of 3 variables (values in 3 columns): day1, day2, day3 for each row. 
dlf$meanHygiene <- rowMeans(cbind(dlf$day1, dlf$day2, dlf$day3) ,  na.rm = TRUE)
```

```{r}

```

