---
title: "Chap 6 exercises"
output: html_notebook
---

```{r}
essayMarks <- read.delim("../Data files/EssayMarks.dat")
essayMarks$grade <- as.factor(essayMarks$grade)
essayMarks$grade_level <- c(1:45)
essayMarks[essayMarks$grade == "First Class",]$grade_level <- 1
essayMarks[essayMarks$grade == "Upper Second Class",]$grade_level <- 2
essayMarks[essayMarks$grade == "Lower Second Class",]$grade_level <- 3
essayMarks[essayMarks$grade == "Third Class",]$grade_level <- 4
str(essayMarks)
```

Both variables are normally distributed by using shapiro-wilk test, we can use Spearman correlation. There is a positive correlation between time spent doing essays and essay numerical grades. We perform positive one-sided correlation test and find that the result is statistically significant. We further create a scatterplot to confirm our findings. We conclude that there is a significant positive relationship between

```{r}
cor(essayMarks$hours, essayMarks$essay)
cor.test(essayMarks$hours, essayMarks$essay, alternative = "greater")
shapiro.test(essayMarks$hours)
shapiro.test(essayMarks$essay)
ggplot(essayMarks, aes(x=hours, y=essay)) + geom_point()
```

We also measure motononic relationship between degree class and essay grades and find that there is a negative assocation. However, the correlation is very weak since the p-value of the negative one-sided correlation test gives us 0.9325 which is statistically insignificant.

```{r}
cor(essayMarks$hours, essayMarks$grade_level, method="kendall")
cor.test(essayMarks$hours, essayMarks$essay, alternative = "less", method="kendall")
```

Since both variables are ordinal, we use Spearman's rho and Kendall's tau instead of Pearson's correlation. We predicted that there is a positive relationship gcse score and first year stat score. Using Spearman's rho and Kendall's tau, we conclude that there is a positive correlation between first year stat grades and gcse grades.The p-value from the one-sided positive correlation test gives us 0.01121 which is statistically significant result.

```{r}
grades <- read.csv("../Data files/grades.csv", header=TRUE)
cor(grades$stats, grades$gcse, method = "spearman")
cor.test(grades$stats, grades$gcse, alternative = "greater", method = "spearman")
cor(grades$stats, grades$gcse, method = "kendall")
cor.test(grades$stats, grades$gcse, alternative = "greater", method = "kendall")
```
