---
title: "Chap 6"
output: html_notebook
---

```{r}
adverts<-c(5,4,4,6,8)
packets<-c(8,9,10,13,15)
advertData<-data.frame(adverts, packets)

ggplot(advertData, aes(x=adverts, y=packets)) + 
    geom_point()
```
```{r}
examData <- read.delim("../Data Files/Exam Anxiety.dat", header = TRUE)
```

```{r}
examData2 <- examData[, c("Exam", "Anxiety", "Revise")]
cor(examData2)
```
```{r}
# confidence interval of correlation between anxiety and exam score. it is a range of values that is likely to capture the true population correlation
cor.test(examData$Anxiety, examData$Exam)
```
```{r}
cor.test(examData$Revise, examData$Exam)
cor.test(examData$Revise, examData$Anxiety)
```
```{r}
cor(examData2) ^ 2
```
```{r}
liarData <- read.delim("../Data files/The Biggest Liar.dat", header = TRUE)
```

```{r}
cor(liarData$Position, liarData$Creativity, method = "spearman")
```
```{r}
cor.test(liarData$Position, liarData$Creativity, alternative = "less",
method = "spearman")
```
```{r}
# good when data points have same rank
cor(liarData$Position, liarData$Creativity, method = "kendall")
```
```{r}

# we know there is a negative correlation so we want to see if there is a signifiant negative association between 2 variables
cor.test(liarData$Position, liarData$Creativity,
alternative = "less", method = "kendall")
```
```{r}
bootTau<-function(liarData,i) cor(liarData$Position[i], liarData$Creativity[i], use = "complete.obs", method = "kendall")
```

```{r}
boot_kendall<-boot(liarData, bootTau, 2000)
boot_kendall
boot.ci(boot_kendall)
```

```{r}
catData = read.csv("../Data Files/pbcorr.csv", header = TRUE)
```

since we code 1 for male and we have positive correlation, we can conclude that male cats spend more time away from home
```{r}
cor(catData$time, catData$gender, method = "pearson")
```
```{r}
cor.test(catData$time, catData$gender)
```
since we code 1 for female and we have negative correlation, we can conclude that female cats spend less time away from home (gender increases (from 0 to 1), time decreases)
```{r}
cor(catData$time, catData$recode, method = "pearson")
```

```{r}
catFrequencies <- table(catData$gender)
prop.table(catFrequencies)
```
biserial correlation assuming that we have some continuum for cat genders. 
```{r}
polyserial(catData$time, catData$gender)
```
when we ignore effect of third variable Revise, the correlation between Exam and Anxiety drops significantly. the exam anxeity can only account for 6% of exam score instead of 19.4 like before
```{r}
pc <- pcor(c("Exam", "Anxiety", "Revise"), var(examData2))
pc
pc ^ 2
pcor.test(pc, 1, 103)
```

```{r}
women_examDat <- examData[examData$Gender == "Female",]
men_examDat <- examData[examData$Gender == "Male",]
```

```{r}
cor(women_examDat$Anxiety, women_examDat$Exam)
cor(men_examDat$Anxiety, men_examDat$Exam)
```

