---
title: "task 2"
output: html_notebook
---

```{r}
library(car)
library(tidyverse)
library(nnet)
library(MASS)
```

```{r}
rm(list = ls())
burnout_data <- read.delim('../../Data files/Burnout.dat', header = TRUE)
glimpse(burnout_data)
burnout_data <-
  burnout_data |>
  mutate(
    burnout = as.factor(if_else(burnout == "Burnt Out", 1, 0))
  )
```

```{r}
burnout_data |>
  filter(burnout == "1") |>
  head(5)
```

We first build a base model using only perceived control and coping style

```{r}
base_model <- glm(data = burnout_data, burnout ~ cope + loc, binomial())
summary(base_model)
exp(base_model$coefficients)
```

```{r}
model_using_all_predictors <- glm(data = burnout_data, burnout ~ cope + loc + teaching + research + pastoral, binomial())
summary(model_using_all_predictors)
exp(model_using_all_predictors$coefficients)
```

```{r}
model_chi <- base_model$deviance - model_using_all_predictors$deviance
chidf <- base_model$df.residual - model_using_all_predictors$df.residual
chisq.prob <- 1 - pchisq(model_chi, chidf)
chisq.prob
```

AIC in model using all predictors is much lower. Also, after looking at the difference in deviance stats, we find that p-value of chisq test is significant. This means that the model using all predictors is a significant improvement over the base model. Coping style and level of perceived control are strongest predictors in our model. teaching score has a negative relationship with burnout. teaching score reduces the odd of having burnout by 89 percent. 
